Boundary Labeling for annotated documents

### 2017-10-01 LaTeX/Paper.tex
Paper: Finished first re-write. Program: No changes.

%TODO: CHANGE!
\setauthor{}{Jakob Klinger}{}{male}
\setadvisor{Pretitle}{Forename Surname}{Posttitle}{male}

@@

\settitle{Boundary Labeling for annotated documents}{Boundary Labeling in annotierten Dokumenten}
\setsubtitle{TestSubtitle}{TestSubtitel}

\setthesis{bachelor}
\setcurriculum{EnglischerCurrName}{DeutscherCurrName} %TODO!

@@

\section{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves. If a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.

In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed outside of the text they are referencing and will be visually connected to the feature they are referencing. (See also \cite{Bekos2007})

The guidelines on how to create an optimal labeling are as follows: the connections should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site. These three easily come into conflict with one another,  as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.

@@

A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} $(V=\{v_1, v_2, ..., v_n\})$ and \emph{Edges} $(E=\{e_1, e_2, ..., e_m\})$. A vertex $v_i (i~\in~\mathbb{N})$ is a featureless object. %Wie  sage ich, dass die Umsetzung der Vertices sehr flexibel ist?
 Each edge $e_i \, (i~\in~\mathbb{N})$ is a relation between two vertices $(E:V\rightarrow  V\times V)$. The vertices forming the edge are considered \emph{adjacent} to each other. Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
 A \emph{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.

\emph{Polylines} are a connected series of \emph{Line segments}. Line segments are straight lines that contain each point between their starting and end point. %Open vs. closed line segments?
\emph{Labels} hold additional information and are represented as boxes containing this information. They are usually placed in the \emph{Label area} which is an area designated to hold labels and is located off to a side so the labels don't obscure anything.
The point or object that this information refers to is called a \emph{Site}. The site and the label are connected via a \emph{Leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area. \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination. \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and einding points in a straight line.
The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.
The location where a leader connects to the label is called the \emph{Port}. It can either be located freely along the label's edge, or restricted to pre-defined positions.

\begin{itemize}
% \item \textbf{Graph:} A Tuple of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
% \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
% \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges. Usually depicted as a point or small shape.
% \item \textbf{Path:} An ordered series of either edges that share one vertex each with the edges listed before and after them, or an ordered series of vertices that have an edge connecting each vertex to the subsequent one. Also known as \textbf{``Graph Walk''}.

 %\item \textbf{Polyline:} A connected series of straight lines.%Ist mehr Information notwendig?

% \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
% \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
% \item \textbf{Leader:} The line or polyline connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
%  \begin{itemize}
 %  \item \textit{S-Leader:} This Leader connects site and label in a straight line.
  % \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
  % \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
  %\end{itemize}
   % For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
% \item \textbf{Port:} The location where the leader connects to the label. It may be restricted to pre-determined locations, like only at the corners.
  \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal. Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
  \item \textbf{Depth-first search:} A searching algorithm that continues along a set of rules as far as possible until either the goal is found, or backtracking is required. A common example is navigating a maze by turning the same way whenever possible.
\end{itemize}
%Mehr Graphenterminologie? (zb: Path) +Polylinien Vertex/Node für Knoten in Polylinien bzw. Graphen -->Eindeutig!

\section{The program}

In our program, the focus was put on keeping the text as readable as possible. This means that leaders are only allowed in the space between words or lines, which can make certain locations for sites unavailable, as there might already too many leaders routed through that area to insert yet another for the new site. 
We also wanted the leaders to be monotonous, which means that they will be one of the shortest possible connection between port and site locations.
Our final priority was to efficiently use the space for the labels, which meant that each label should initially be placed as close to the available area's border as possible.
While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders.%''Final Priority'' klingt komisch -> Ersatz finden!

%Port-Location erwähnen!
\subsection{Algorithm}%''Underlying Algorithm''?
To easily create leaders that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing vertices above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point. We also decided to give the graph's edges a capacity of 1, which means that each leader will always use edges which haven't already be taken by other leaders. While this causes some sites to become un-routable due to another leader passing right through their starting vertex, it makes visualizing the leaders much easier.

Our first approach was going through the labels in the order they appear in the text and placing each as far up as possible, to maximize the remaining space other labels could use. We also opted to use fixed ports located in the top left corner of the annotation, as this allowed us to unambiguously place any annotation by knowing either their port's location, wich is also the leader's ending point.
However, since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing, we then decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position. As this allowed us to place the labels more freely, we adjusted the label's positions after all labels were placed to reduce the distance that needed to be covered in the buffer area.

\subsection{Implementation}
The program was written in Java, using only JGraphT\cite{JGraphT} as additional library. Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas, as it was easy to extract measurements at this point. Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.

As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent). This additional information is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further above if possible, and vertices to the right otherwise. After the algorithm terminates, it returns  a vertex-based graph walk describing the route from the site to the border of the text area alongside information about where to draw the OPO-segment in the buffer zone if successful, or a walk containing only the starting vertex, if no valid path was found. Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be easily drawn.
%TODO: Schönheitsfeatures, etc.
%TODO: Code genauer durchlesen und erweitern!

>>>>

%TODO: CHANGE!
\setauthor{}{Jakob Klinger}{}{male}
\setadvisor{}{}{}{}%{Pretitle}{Forename Surname}{Posttitle}{male}
%Ass.Prof. Dipl.-Inform. Dr.rer.nat. Martin Nöllenburg
% University Assistant, M.Sc., B.Sc. Fabian Klute

@@

\settitle{Boundary Labeling for annotated documents}{Boundary Labeling in annotierten Dokumenten}
\setsubtitle{}{}%{TestSubtitleENG}{TestSubtitelGER}

\setthesis{bachelor}
\setcurriculum{Bachelor's programme Software \& Information Engineering}{Bachelorstudium Software \& Information Engineering} %TODO!

@@

\chapter{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
Annotating a document is usually solved by adding footnotes in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves. If a more obvious connection between the text and the referenced content is required, the reference placed to the side of the document and visibly connected to the text by drawing a straight line or a more complex path between them.

In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing. (See also \cite{Bekos2007})

The guidelines on how to create suitable labelings are as follows: the connections should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs where. These three criteria easily come into conflict with one another, as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.

@@

A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} $(V=\{v_1, v_2, ..., v_n\})$ and \emph{Edges} $(E=\{e_1, e_2, ..., e_m\})$. A vertex $v_i (i~\in~\mathbb{N})$ is a featureless object. %Wie  sage ich, dass die Umsetzung der Vertices sehr flexibel ist?
 Each edge $e_i \, (i~\in~\mathbb{N})$ is a relation between two vertices $(E:V\rightarrow  V\times V)$. The vertices forming the edge are considered \emph{adjacent} to each other. Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
 A \emph{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
 \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v_s \in V$ and explores the graph by traversing its edges as far as possible before backtracking. In our algorithm, the goal is to reach a target vertex $v_t \in V$, at which point the algorithm terminates and returns the route taken from $v_s$ to $v_t$.
% A search algorithm is \emph{monotonous} if, for each vertex $v$ and each successor $v'$ generated by the algorithm, the estimated cost of reaching the goal from $v$ is not greater than the cost of reaching $v'$ plus the estimated cost of reaching the goal from $v'$.

\emph{Polylines} are a connected series of \emph{Line segments}. Line segments are straight lines that contain each point between their starting and end point. %Open vs. closed line segments?
\emph{Labels} hold additional information and are represented as boxes containing this information. They are usually placed in the \emph{Label area} which is an area designated to hold labels and is located off to a side so the labels don't obscure anything.
The point or object that this information refers to is called a \emph{Site}. The site and the label are connected via a \emph{Leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area. \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination. \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and einding points in a straight line.
The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.
The location where a leader connects to the label is called the \emph{Port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity

% \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal. Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
%+Node für Knoten in Polylinien?

\chapter{The Algorithm}

In our algorithm, the focus was put on keeping the text as readable as possible. This means that leaders aren't allowed to pass through words, which was simplified by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label in at least one dimension.%Wie drücke ich das als Formel aus?
We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of possibly increased leader length.

%\subsection{Algorithm}%''Underlying Algorithm''?
To easily create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
As each vertex represents a physical location, they will have co-ordinates associated with them, and the vertices representing the sites will hold additional information regarding its leader and label. 
The graph is constructed by placing vertices in the space between lines next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners. %Insert picture
For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point. After placing all vertices, we created edges between each vertex and the closest horizontal neighbour to both sides, and between nodes that are located exactly above or below each other, and exactly one line apart.

In our labeling algorithm, we decided to work through the labels in the order they appear in the text, and placing each as far up as possible, skipping any label that would've required us to move a lower line. We also opted to use fixed ports located in the top left corner of each label, as this allowed us to unabiguously place annotations by only knowing their port location, which is equal to their leader's ending point. To not restrict the label's placement by the line spacing, we left a buffer zone between the text area and the label area, which allows to place labels even further above to optimally use the available space.
The routing was implemented using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex, prioritizimg locations above whenever possible. After the algorithm terminates, it returns a path from the site to the right text border and information on where to draw the OPO-segment that connects the text area with the label's port, if successful.

\section{Implementation}
The program was written in Java 1.8.0u40, using JGraphT1.0.1\cite{JGraphT} as graph library. Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point. Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.

%As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
%The additional information stored in the vertices is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further up if possible, and vertices to the right otherwise. After the algorithm terminates, it returns  a vertex-based graph walk describing the route from the site to the border of the text area alongside information about where to draw the OPO-segment in the buffer zone if successful, or a walk containing only the starting vertex, if no valid path was found. Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be drawn.
%TODO: Schönheitsfeatures, etc.
%TODO: Code genauer durchlesen und erweitern!
